Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-14 doc/mc-paper.tex
Ring must be commutative

\newcommand{\indegree}{\textrm{indegree}}
\newcommand{\outdegree}{\textrm{outdegree}}
\newcommand{\validate}{\textsf{validate}}
\newcommand{\prov}{\textrm{Prov}}

@@

@@

Fichte, Hecher, and Roland~\cite{fichte:sat:2021} devised the MICE
proof framework for model counting programs.  Their proof rules are
based on the algorithmic steps commonly used by model counting
programs, and they present a series of lemmas and a theorem stating
that their checking framework is sound and complete.  They also
present experimental results for cases where they modify existing
model counters to also generate proof traces, and where they convert
d-DNNF graphs into a checkable form.  We compare the performance of
our tool chain to theirs as part of our experimental evaluation in
Section~\ref{sect:experimental}.  Our experience with developing a
formally verified proof system has shown us that, even for the well
understood framework of extended resolution, ensuring that the system
is sound can be challenging.

@@

\begin{definition}[Ring Evaluation Problem (REP)]
  For ring $\ring$, a {\em ring labeling} associates a value $w(x) \in \dset$ with
  every variable $x \in \varset$.  We then define $w(\obar{x})$ to equal $\mulident-w(x)$.

@@

Many important queries about Boolean formulas can be
expressed as ring evaluation problems.  The 
{\em model counting} problem for formula $\phi$ requires determining $|\modelset(\phi)|$.
It can be cast as a REP by having $\radd$ and
$\rmul$ be addition and multiplication over rational numbers and using
label $w(x) = 1/2$ for every variable $x$.  
Ring evaluation of formula $\phi$ will then compute the {\em density} of
the formula, i.e., the fraction of all possible total assignments that are
models.  For $n = |\varset|$, scaling the density by $2^n$ will then
yield the number of models.

@@

\subsection{Syntax}

@@

\section{Generating CRAT from d-DNNF}

>>>>

\newcommand{\indegree}{\textrm{indegree}}
\newcommand{\outdegree}{\textrm{outdegree}}
\newcommand{\validate}{\textsf{validate}}
\newcommand{\prov}{\textrm{Prov}}
\newcommand{\inputformula}{\phi_I}
\newcommand{\pogformula}{\phi_P}

@@

% Lean code:
\usepackage{listings}
\definecolor{keywordcolor}{rgb}{0.7, 0.1, 0.1}   % red
\definecolor{tacticcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{commentcolor}{rgb}{0.4, 0.4, 0.4}   % grey
\definecolor{symbolcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{sortcolor}{rgb}{0.1, 0.5, 0.1}      % green
\definecolor{attributecolor}{rgb}{0.7, 0.1, 0.1} % red
\def\lstlanguagefiles{lstlean.tex}
% set default language
\lstset{language=lean, xleftmargin=1em}
\lstset{backgroundcolor=\color{white}}

@@

Fichte, Hecher, and Roland~\cite{fichte:sat:2022} devised the MICE
proof framework for model counting programs.  Their proof rules are
based on the algorithmic steps commonly used by model counting
programs, and they present a series of lemmas and a theorem stating
that their checking framework is sound and complete.  They also
present experimental results for cases where they modify existing
model counters to also generate proof traces, and where they convert
d-DNNF graphs into a checkable form.  We compare the performance of
our tool chain to theirs as part of our experimental evaluation in
Section~\ref{sect:experimental}.  Our experience with developing a
formally verified proof system has shown us that, even for the well
understood framework of extended resolution, ensuring that the system
is sound can be challenging.

@@

\begin{definition}[Ring Evaluation Problem]
  For commutative ring $\ring$, a {\em ring labeling} associates a value $w(x) \in \dset$ with
  every variable $x \in \varset$.  We then define $w(\obar{x})$ to equal $\mulident-w(x)$.

@@

Many important queries about Boolean formulas can be
expressed as ring evaluation problems.  The
{\em model counting} problem for formula $\phi$ requires determining $|\modelset(\phi)|$.
It can be cast as a ring evaluation problem by having $\radd$ and
$\rmul$ be addition and multiplication over rational numbers and using
label $w(x) = 1/2$ for every variable $x$.
Ring evaluation of formula $\phi$ will then compute the {\em density} of
the formula, i.e., the fraction of all possible total assignments that are
models.  For $n = |\varset|$, scaling the density by $2^n$ will then
yield the number of models.

@@

\subsection{Syntax}
\label{subsection:syntax}

@@

\section{Generating CRAT from d-DNNF}
\label{section:generating:crat}
